# PRACTICAL MACHINE LEARNING PROJECT
## Executive Summary
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively.In this project,goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants and predict the manner in which they did the exercise.

##Data
The training data for this project are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
##Method
1. To clean the data
2. To split the training dataset into sub-test and sub-train sets
3. To build the model on sub-train set and test on sub-test set
4. Validate on testing data set
5. Estimate the errors
##Analysis
```{r echo=TRUE, results='hide'}
# loading the required libraries
library(caret)
library(randomForest)
```
```{r, echo=TRUE}
# reading the training and test datasets
training<-read.csv("C:/Users/Praveen Gorthy/Desktop/assg/specdata/pml-training.csv")
testing<-read.csv("C:/Users/Praveen Gorthy/Desktop/assg/specdata/pml-testing.csv")
table(training$classe)
```
Setting the seed and dividing the training data into sub-training and sub-testing data sets
```{r echo=TRUE}
set.seed(333111)
shorttrain <- createDataPartition(y=training$classe,p=0.6,list=F)
trainsub <- training[shorttrain,]
testsub <- training[-shorttrain,]
```
##Cleaning the data
Removing the near zero values  from training data set
```{r echo=TRUE}
nearzero1 <- nearZeroVar(trainsub)
training<- trainsub[,-nearzero1]
```
Columns containing many NA's were removed from training dataset
```{r,echo=TRUE}
trainnew <- training[, which(as.numeric(colSums(is.na(training)))==0)]
```
The datasets have been reduced by removing descriptive columns from training and testing datasets retaining only the useful predictor variables
```{r echo=TRUE}
trainnew <- trainnew[,-(1:7)]
```
##Regression analysis
Fitting the model using the random forest method over the sub train data set
```{r echo=TRUE}
modfit<- randomForest(classe~.,data=trainnew,importance=T)
```
##Predicting 
The model thus buld is used to predict over the sub-test data set 
```{r echo=TRUE}
predictTestsub<-predict(modfit,testsub)

confusionMatrix(predictTestsub,testsub$classe)
```
The out of sample accuracy is 99.42% , the error is about 0.58%. 
##Testing 
Testing the prediction on the actual testing data set provided
the results are:
```{r echo=TRUE}
predictTest<-predict(modfit,testing)
predictTest
```
## Prediction submmission
```{r echo=True} 
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
```

